---
title: "US Ski resorts - Leaflet Map"
author: "Iñigo Fernández del Amo"
date: "1/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Developing Data Products - Week 2 Assignment

## Executive summary

The purpose of this work is to display two maps (unclustered and clustered) showing the US ski resorts with markers colored by the level of difficulty of their slopes.

The data utilized for this work is publicly available and can be found [**here**](!https://gist.githubusercontent.com/Ewiseman/b251e5eaf70ca52a4b9b10dce9e635a4/raw/9f0100fe14169a058c451380edca3bda24d3f673/ski_resort_stats.csv)

```{r}
# Libraries used for this work
library(knitr)
library(leaflet)
library(fontawesome)
library(dplyr)
# US ski resorts data url
skiURL <- "https://gist.githubusercontent.com/Ewiseman/b251e5eaf70ca52a4b9b10dce9e635a4/raw/9f0100fe14169a058c451380edca3bda24d3f673/ski_resort_stats.csv"
```

## US Ski resorts - Unclustered map

The first map shows all the ski resorts in the US through markers colored by their slopes' most common level of difficulty and popping-up the ski resort's name.

```{r}
# Download ski data from public URL omitting rows with NA values and pre-process
skiData <- na.omit(read.csv(url(skiURL)))
# Identify the highest percentage of slope types (green, blue or black)
skiData <- skiData %>% 
  mutate (slope_trend = 
            if_else(green_percent >= blue_percent, 
                    if_else(green_percent >= black_percent, 'green', 'black'), 
                    if_else(blue_percent >= black_percent, 'blue', 'black')))
# Create map using ad-hoc icons from font awesome: https://github.com/rstudio/fontawesome
# Markers will be colored according to slope_trend
# Markers will have as pop-up the ski resort_name
skiMap <- leaflet(data = skiData) %>% 
  addTiles() %>% 
  addAwesomeMarkers(~lon, ~lat,
                    icon = makeAwesomeIcon(text = fa(name = "skiing",
                                                     fill = "white"), 
                                           markerColor = ~slope_trend),
                    popup = ~resort_name)
skiMap
```

## US Ski resorts - Clustered map

The second map shows all the US ski resorts similarly to the first, although this time they are clustered and there is a legend to identify the color with the slopes' level of difficulty.

```{r}
# Legend will be added to show meaning of markers' color
skiLegend <- c("#72AF26", "#36A4D5", "#303030")
# Create secondary map using cluster options to simplify map
skiMapCluster <- leaflet(data = skiData) %>% 
  addTiles() %>% 
  addAwesomeMarkers(~lon, ~lat,
                    icon = makeAwesomeIcon(text = fa(name = "skiing",
                                                     fill = "white"), 
                                           markerColor = ~slope_trend),
                    popup = ~resort_name,
                    options = markerOptions(trend = ~slope_trend),
                    clusterOptions = markerClusterOptions()) %>%
  addLegend(position = "bottomright", title = "Resort slopes' difficulty",
            labels = c("Easy", "Medium", "Hard"), 
            colors = skiLegend)
skiMapCluster
```

